A Short and Mechanized Consistency Proof for Dependent Type Theory

### 2024-02-11 paper/paper.tex
Minor comments

Figure~\ref{fig:par} shows the definition of the parallel reduction
relation, which takes the form $[[a => b]]$. We use $[[a =>+ b]]$ to
represent its transitive closure, which in turn allows us to define
the coherence relation $[[a <=> b]]$.
We say that two terms $[[a]]$
and $[[b]]$ are coherent if they can eventually reduce to some common
term $[[c]]$ through parallel reduction. The symmetric notation of
coherence suggests that it is an equivalence relation, and in fact, we
will use coherence as the equational theory for our type conversion
rule.

We sketch out some key properties about coherence without giving their
proofs. Our technique for establishing those results is based on
\citet{takahashi-parallel-reduction} and a modern exposition can be
found in \citet{plfa22.08}.

@@

We note that a more conventional presentation of
\rref{T-Conv} would instead use full beta reduction as the base for
the definition of coherence. However, since full beta reduction
doesn't satisfy the diamond property, one typically needs parallel
reduction as an auxilliary definition to derive the confluence of full
beta reduction. Our formulation of \lang through parallel reduction
is slightly more economical.

>>>>

Figure~\ref{fig:par} shows the definition of the parallel reduction
relation, which takes the form $[[a => b]]$. We use $[[a =>+ b]]$ to
represent its transitive and reflexive closure, which in turn allows us to define
the coherence relation $[[a <=> b]]$.
We say that two terms $[[a]]$
and $[[b]]$ are coherent if they can eventually reduce to some common
term $[[c]]$ through parallel reduction. The symmetric notation of
coherence suggests that it is an equivalence relation, and in fact, we
will use coherence as the equational theory for our type conversion
rule.

We sketch out some key properties about coherence without giving their
proofs. Our technique for establishing those results is based on
\citet{takahashi-parallel-reduction} and a modern exposition of the
same technique can be found in \citet{plfa22.08}.

TODO: remove the lengthy discussion below

@@

Figure~\ref{fig:typing} gives the full typing rules for
\lang{}. \Rref{T-Conv} uses the definition of coherence from earlier
for type conversion. The use of an untyped relation for type
conversion makes our formulation slightly different from languages
such as MLTT, where the judgmental equality typically takes the form
TODO, where terms ... and ... are known to be well-typed. The
equivalence of such systems and a system that uses untyped equality
are explored in detail in ...

Without fancy eta laws, it is easy to embed a typed language into an
untyped language.

% We note that a more conventional presentation of
% \rref{T-Conv} would instead use full beta reduction as the base for
% the definition of coherence. However, since full beta reduction
% doesn't satisfy the diamond property, one typically needs parallel
% reduction as an auxilliary definition to derive the confluence of full
% beta reduction. Our formulation of \lang through parallel reduction
% is slightly more economical.
